---
title: "ANOVA Interactive"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(car)# for Anova
library(ggplot2)# for plots
library(plotly)# plots
library(datasets)
```
ANOVA (short of Analysis of Variance) is a statistical tool for determining differences/Variances between two or more groups of independent variables.
One way ANOVA checks for one independent variable whereas Two way ANOVA checks for two independent variables.
They can also be checked for interaction effect i.e. The presence of one causes changes in the other.
Further t test is performed to find which level of independent variable shows variation.
Independent variable and Dependent variable are Categorical and Metric respectively.
```{r,echo=FALSE}
cat(sprintf("\n "))

#inputpanel
sidebarPanel(
  inputPanel(
    checkboxInput("Choice","Check to choose your own file",value=FALSE),
    fileInput("file","Upload a file in *.csv format with headers",placeholder = "No file selected"),
    uiOutput("Var1"),
    uiOutput("Var2"),
    uiOutput("Var3"),
    selectInput("way", " select the anova technique to be used", choices= c("one way","two way")),
    uiOutput("int"),
    downloadButton("dataset"," Download dataset"),
    downloadButton("plot","Download Plot")
    
  )
)


  #selection ofvariable 1
output$Var1<- renderUI({
  if (input$Choice==FALSE) {
  data("iris")
    
  data= iris
    data=select_if(data,is.numeric)
  selectInput("var1"," Select the dependent variable",choices = colnames(data))
  }
  else {
  
  file1=input$file
  if (is.null(file1)){return()}
  
  data = read.table(file =  file1$datapath,sep = ",",header = TRUE)
  if (is.null(data())){return()}
  
  data=select_if(data,is.numeric)
  selectInput("var1"," Select the dependent variable",choices = colnames(data))
  }
})

#selection of variable 2
output$Var2<- renderUI({
  if (input$Choice==FALSE) {
  
  data("iris")
    
  data= iris
  data=select_if(data,is.factor)
  selectInput("var2","Select the independent variable 1",choices = colnames(data))
  }
  
else{
  file1=input$file
  if (is.null(file1)){return()}
  
  data = read.table(file =  file1$datapath,sep = ",",header = TRUE)
  if (is.null(data())){return()}
  
  data=select_if(data,is.factor)
  selectInput("var2","Select the independent variable 1",choices = colnames(data))
   }
  })


#selection of third variable for two way
output$Var3<- renderUI({
  
  if (input$way=="two way") {
    if (input$Choice==FALSE) {
  
  data("iris")
    
  data= iris
    data=select_if(data,is.factor)
    data=select(data,-c(input$var2))
    selectInput("var3","Select the independent variable 2",choices = colnames(data))
    }
   else{
  file1=input$file
  if (is.null(file1)){return()}
  
  data = read.table(file =  file1$datapath,sep = ",",header = TRUE)
  if (is.null(data())){return()}
  
  
    data=select_if(data,is.factor)
    data=select(data,-c(input$var2))
    selectInput("var3","Select the independent variable 2",choices = colnames(data))
  }
   }
})



#selection of whether or not to check for interaction effect
output$int<-renderUI({
  
  if (input$way=="two way") {
    selectInput("inter","Please select whether to check for interaction or not",choices = c("no","yes"))
  }
  
})


# tabs
mainPanel(
  
  tabsetPanel(
    tabPanel("Summary",verbatimTextOutput("summ")),
    tabPanel("Anova",verbatimTextOutput("ANOVA")),
    tabPanel("Post Hoc(for single variables) ",verbatimTextOutput("PH")),
    tabPanel("Visualization",plotlyOutput("Plot"))
  ),
  h6("Designed and Developed by : Ms. Revathi Kumar", tags$img(src ="R.JPG", height= 200, width=200)),
  h6("Mentored by  :", tags$img(src ="K.JPG", height= 300, width=300))

  
 )


#for the summary of the data

output$summ<- renderPrint({
  if (input$Choice==FALSE) {
  
  data("iris")
    
  data= iris
  
  str(data)
  
  summary(data)
  
  cat(sprintf("\n The summary of the dataset is as follows:\n"))
  print(summary(data))
  }
  else{
    
  file1=input$file
  if (is.null(file1)){return()}
  
  data = read.table(file =  file1$datapath,sep = ",",header = TRUE)
  if (is.null(data())){return()}
  
  str(data)
  
  summary(data)
  
  cat(sprintf("\n The summary of the dataset is as follows:\n"))
  print(summary(data))
  
  }
})


#for ANOVA test
output$ANOVA<- renderPrint({
  
  if (input$Choice==FALSE) {
    
  data("iris")
    
  data= iris
 
  if (input$way=="one way") {
    cat(sprintf("\n Null Hypothesis : means of %s is uniform across all groups of %s\n",input$var1,input$var2))
    cat(sprintf("\n Alternate Hypothesis : means of %s is not uniform across all groups of %s\n",input$var1,input$var2))
    mod= paste(input$var1,input$var2,sep = "~")
    
    model=lm(formula = as.formula(mod),data = data)
    a<- aov(model)
    
    summary(a)
    
  }
  
  
  else {
    
    if (input$inter=="no") {
      
      cat(sprintf("\n Hypothesis 1\n"))
      cat(sprintf("\n Null Hypothesis : means of %s is uniform across all groups of %s\n",input$var1,input$var2))
      cat(sprintf("\n Alternate Hypothesis : means of %s is not uniform across all groups of %s\n",input$var1,input$var2))
      
      cat(sprintf("\n Hypothesis 2\n"))  
      cat(sprintf("\n Null Hypothesis : means of %s is uniform across all groups of %s\n",input$var1,input$var3))
      cat(sprintf("\n Alternate Hypothesis : means of %s is not uniform across all groups of %s\n",input$var1,input$var3))
      mo=paste(input$var2,input$var3,sep = "+")
      mod1= paste(input$var1,mo,sep = "~")
      
      model=lm(formula = as.formula(mod1),data = data)
      a<- Anova(model)
      
    }

    else{
      
      cat(sprintf("\n Hypothesis 1\n"))
      cat(sprintf("\n Null Hypothesis : means of %s is uniform across all groups of %s\n",input$var1,input$var2))
      cat(sprintf("\n Alternate Hypothesis : means of %s is not uniform across all groups of %s\n",input$var1,input$var2))
      
      cat(sprintf("\n Hypothesis 2\n"))  
      cat(sprintf("\n Null Hypothesis : means of %s is uniform across all groups of %s\n",input$var1,input$var3))
      cat(sprintf("\n Alternate Hypothesis : means of %s is not uniform across all groups of %s\n",input$var1,input$var3))
      cat(sprintf("\n Hypothesis 3\n"))  
      cat(sprintf("\n Null Hypothesis : there is no interaction between %s and %s",input$var2,input$var3))
      cat(sprintf("\n Alternate Hypothesis : there is interaction between %s and %s\n",input$var2,input$var3))
      m=paste(input$var2,input$var3,sep = "*")
      mo=paste(input$var2,input$var3,m,sep = "+")
      mod1= paste(input$var1,mo,sep = "~")
      
      model=lm(formula = as.formula(mod1),data = data)
      a<- Anova(model)
    }
    
    print(a)
    
  }
  }
  else {
    
   file1=input$file
  if (is.null(file1)){return()}
  
  data = read.table(file =  file1$datapath,sep = ",",header = TRUE)
  if (is.null(data())){return()}
 
  if (input$way=="one way") {
    cat(sprintf("\n Null Hypothesis : means of %s is uniform across all groups of %s\n",input$var1,input$var2))
    cat(sprintf("\n Alternate Hypothesis : means of %s is not uniform across all groups of %s\n",input$var1,input$var2))
    mod= paste(input$var1,input$var2,sep = "~")
    
    model=lm(formula = as.formula(mod),data = data)
    a<- aov(model)
    
    summary(a)
    
  }
  
  
  else {
    
    if (input$inter=="no") {
      
      cat(sprintf("\n Hypothesis 1\n"))
      cat(sprintf("\n Null Hypothesis : means of %s is uniform across all groups of %s\n",input$var1,input$var2))
      cat(sprintf("\n Alternate Hypothesis : means of %s is not uniform across all groups of %s\n",input$var1,input$var2))
      
      cat(sprintf("\n Hypothesis 2\n"))  
      cat(sprintf("\n Null Hypothesis : means of %s is uniform across all groups of %s\n",input$var1,input$var3))
      cat(sprintf("\n Alternate Hypothesis : means of %s is not uniform across all groups of %s\n",input$var1,input$var3))
      mo=paste(input$var2,input$var3,sep = "+")
      mod1= paste(input$var1,mo,sep = "~")
      
      model=lm(formula = as.formula(mod1),data = data)
      a<- Anova(model)
      
    }

    else{
      
      cat(sprintf("\n Hypothesis 1\n"))
      cat(sprintf("\n Null Hypothesis : means of %s is uniform across all groups of %s\n",input$var1,input$var2))
      cat(sprintf("\n Alternate Hypothesis : means of %s is not uniform across all groups of %s\n",input$var1,input$var2))
      
      cat(sprintf("\n Hypothesis 2\n"))  
      cat(sprintf("\n Null Hypothesis : means of %s is uniform across all groups of %s\n",input$var1,input$var3))
      cat(sprintf("\n Alternate Hypothesis : means of %s is not uniform across all groups of %s\n",input$var1,input$var3))
      cat(sprintf("\n Hypothesis 3\n"))  
      cat(sprintf("\n Null Hypothesis : there is no interaction between %s and %s",input$var2,input$var3))
      cat(sprintf("\n Alternate Hypothesis : there is interaction between %s and %s\n",input$var2,input$var3))
      m=paste(input$var2,input$var3,sep = "*")
      mo=paste(input$var2,input$var3,m,sep = "+")
      mod1= paste(input$var1,mo,sep = "~")
      
      model=lm(formula = as.formula(mod1),data = data)
      a<- Anova(model)
    }
    
    print(a)
    
  }
  }
})
  



#for Post Hoc test
output$PH<- renderPrint({
  
  if(input$Choice==FALSE){
    data("iris")
    
  data= iris
    
  mod= paste(input$var1,input$var2,sep = "~")
  
  model=lm(formula = as.formula(mod),data = data)
  a<- aov(model)
  
  TukeyHSD(a)
  }
  else{
     file1=input$file
  if (is.null(file1)){return()}
  
  data = read.table(file =  file1$datapath,sep = ",",header = TRUE)
  if (is.null(data())){return()}
  
  mod= paste(input$var1,input$var2,sep = "~")
  
  model=lm(formula = as.formula(mod),data = data)
  a<- aov(model)
  
  TukeyHSD(a)
  }
  
})


#plot 
output$Plot <- renderPlotly({
  
  if(input$Choice==FALSE){
    data("iris")
    
  data= iris
  if(input$way=="one way"){
    
    
    b<-ggplot(data = data, aes_string(x=  input$var2,y= input$var1, fill = input$var2 ))+labs(y=input$var1,x=input$var2)+ geom_boxplot()
    
    b<- ggplotly(b)
    
    print(b)}
  
  else
  {
    
    
    w= aggregate(data[,input$var1],by=list(k= data[,input$var2],l= data[,input$var3]),mean)
    
    b<-ggplot(data= w, aes_string(x= w$k, y= w$x,fill= w$l ))+labs(y=input$var1,x=input$var2,fill= input$var3 )+theme(legend.position = "top")+geom_line(aes_string(group= w$l))+ geom_point() 
    
    b<- ggplotly(b)
    
    print(b)
  }
  }
  else{
     file1=input$file
  if (is.null(file1)){return()}
  
  data = read.table(file =  file1$datapath,sep = ",",header = TRUE)
  if (is.null(data())){return()}
  
    if(input$way=="one way"){
    
    
    b<-ggplot(data = data, aes_string(x=  input$var2,y= input$var1, fill = input$var2 ))+labs(y=input$var1,x=input$var2)+ geom_boxplot()
    
    b<- ggplotly(b)
    
    print(b)}
  
  else
  {
    
    
    w= aggregate(data[,input$var1],by=list(k= data[,input$var2],l= data[,input$var3]),mean)
    
    b<-ggplot(data= w, aes_string(x= w$k, y= w$x,fill= w$l ))+labs(y=input$var1,x=input$var2,fill= input$var3 )+theme(legend.position = "top")+geom_line(aes_string(group= w$l))+ geom_point() 
    
    b<- ggplotly(b)
    
    print(b)
  }
  }
  
})

#for downloading the dataset
datasetInput1<- reactive({
  
  if(input$Choice==FALSE){
    data("iris")
    
  data= iris
  data1<- data.frame(data)
  }
  else
  {
    file1=input$file
  if (is.null(file1)){return()}
  
  data = read.table(file =  file1$datapath,sep = ",",header = TRUE)
  if (is.null(data())){return()}
  data1<-data.frame(data)
  }
  
})

output$dataset<-downloadHandler(
  filename = function(){
    paste("dataset",".csv",sep = "")
  },
  content = function(file){write.csv(datasetInput1(),file,row.names = FALSE)}
)

plotInput1<-reactive({
  
  if (input$Choice==FALSE) {
    data("iris")
    
  data= iris
  }
    else{
    file1=input$file
  if (is.null(file1)){return()}
  
  data = read.table(file =  file1$datapath,sep = ",",header = TRUE)
  if (is.null(data())){return()}
  }
  
  if(input$way=="one way"){
    
    
    b<-ggplot(data = data, aes_string(x=  input$var2,y= input$var1, fill = input$var2 ))+labs(y=input$var1,x=input$var2)+ geom_boxplot() 
  
    
    print(b)
    
    }
  
  else
  {
    
    w= aggregate(data[,input$var1],by=list(k= data[,input$var2],l= data[,input$var3]),mean)
    
    b<-ggplot(data= w, aes_string(x= w$k, y= w$x,fill= w$l ))+labs(y=input$var1,x=input$var2,fill= input$var3 )+theme(legend.position = "top")+geom_line(aes_string(group= w$l))+ geom_point() 
  
    
    print(b)
  }
  
})

output$plot<-downloadHandler(
  filename = function() { paste("PLOT", '.png', sep='') },
  content = function(file) {
    ggsave(file,plotInput1())
    })









```
